---
title: "Modeling NBA Number of Games Played by Players"
author: "Carson Buttars"
date: ""
output: pdf_document
---

```{r setup, include=FALSE}
library(DBI)
library(tidyverse)
library(lubridate)
library(devtools)
load_all('lib/advanced.stats/')
conn = dbConnect(RSQLite::SQLite(), "../data/bball_db")

logit <- function(prob) {
	output <- log(prob / (1 - prob))
	return(output)
}
expit <- function(odds) {
	output <- exp(odds) / (1 + exp(odds))
	return(output)
}
```

# Introduction

In an effort to make a simple and useful tool for predicting performance among NBA players,
I start with this analysis of the number of games played in a season given past performance.
The data to be used comes from the web-scraping script found in this same repository,
and it comprised of 20 years worth of game-by-game and quarter-by-quarter data of individual and team data.
There are some obvious mistakes in the data,
and there are likely some that are more subtle.
Furthermore, I have made some approximations to ease some calculations.
Regardless, if the data is not accurate at any point,
those mistakes are likely to be outweighed by the sheer quantity of observations.

As a premise for this analysis,
I am only analyzing players who would be reasonably likely to be picked for a roughly-standard fantasy basketball team as this is my main goal.
The criteria will be as follows: Either a player..
- Played at least 1000 minutes in a season
- Or they were injured for some portion of games and averaged over 20 minutes per game

```{r data_cleanup, include=FALSE}
birthdays <- dbReadTable(conn, "player_info") %>%
	as_tibble() %>%
	mutate(Birthday = mdy(Birthday)) %>%
	select(Player_ID, Birthday)

season_start <- dbReadTable(conn, "game_info") %>%
	as_tibble() %>%
	mutate(Date = mdy(Date)) %>%
	group_by(Season) %>%
	summarize(Start_Date = first(Date)) %>%
	ungroup() %>%
	select(Season, Start_Date)

team_games <- dbReadTable(conn, "team_games") %>%
	rename(MP = Total_minutes) %>%
	as_tibble()

team_and_opponents <- team_games %>%
	inner_join(team_games, by = c("Game_ID" = "Game_ID",
				      "Opponent_ID" = "Team_ID",
				      "Team_ID" = "Opponent_ID",
				      "Season" = "Season"),
		   suffix = c("_team", "_opponent")) %>%
	mutate(Pace_team = pace(MP_team, Possessions_team, Possessions_opponent),
	       Pace_opponent = pace(MP_opponent, Possessions_opponent, Possessions_team))

league <- team_and_opponents %>%
	group_by(Season) %>%
	summarize(league_Games = n(),
		  league_MP = sum(MP_team),
		  league_Possessions_team = sum(Possessions_team),
		  league_Threes_team = sum(Threes_team),
		  league_Three_Attempts_team = sum(Three_Attempts_team),
		  league_Twos_team = sum(Twos_team),
		  league_Two_Attempts_team = sum(Two_Attempts_team),
		  league_Freethrows_team = sum(Freethrows_team),
		  league_Freethrow_Attempts_team = sum(Freethrow_Attempts_team),
		  league_Offensive_Rebounds_team = sum(Offensive_Rebounds_team),
		  league_Deffensive_Rebounds_team = sum(Deffensive_Rebounds_team),
		  league_Assists_team = sum(Assists_team),
		  league_Steals_team = sum(Steals_team),
		  league_Blocks_team = sum(Blocks_team),
		  league_Turnovers_team = sum(Turnovers_team),
		  league_Fouls_team = sum(Fouls_team),
		  league_Points_team = sum(Points_team),
		  league_Win_team = mean(Win_team),
		  league_Home_team = mean(Home_team),
		  league_MP = sum(MP_opponent),
		  league_Possessions_opponent = sum(Possessions_opponent),
		  league_Threes_opponent = sum(Threes_opponent),
		  league_Three_Attempts_opponent = sum(Three_Attempts_opponent),
		  league_Twos_opponent = sum(Twos_opponent),
		  league_Two_Attempts_opponent = sum(Two_Attempts_opponent),
		  league_Freethrows_opponent = sum(Freethrows_opponent),
		  league_Freethrow_Attempts_opponent = sum(Freethrow_Attempts_opponent),
		  league_Offensive_Rebounds_opponent = sum(Offensive_Rebounds_opponent),
		  league_Deffensive_Rebounds_opponent = sum(Deffensive_Rebounds_opponent),
		  league_Assists_opponent = sum(Assists_opponent),
		  league_Steals_opponent = sum(Steals_opponent),
		  league_Blocks_opponent = sum(Blocks_opponent),
		  league_Turnovers_opponent = sum(Turnovers_opponent),
		  league_Fouls_opponent = sum(Fouls_opponent),
		  league_Points_opponent = sum(Points_opponent),
		  league_Win_opponent = mean(Win_opponent),
		  league_Home_opponent = mean(Home_opponent),
		  league_Pts_per_game = mean(Points_team)
	) %>%
	ungroup() %>%
	mutate(league_Pace = pace(league_MP, league_Possessions_team, league_Possessions_opponent),
	       league_Pts_per_Poss = league_Points_team / league_Possessions_team)

player_season_totals <- dbReadTable(conn, "player_games") %>%
	as_tibble() %>%
	inner_join(team_and_opponents, by = c("Game_ID" = "Game_ID",
					      "Team_ID" = "Team_ID",
					      "Opponent_ID" = "Opponent_ID",
					      "Season" = "Season")) %>%
	mutate(Played = Seconds > 0,
	       Usage = usage(Seconds / 60,
			     Two_Attempts + Three_Attempts,
			     Freethrow_Attempts,
			     Turnovers,
			     MP_team,
			     Two_Attempts_team + Three_Attempts_team,
			     Freethrow_Attempts_team,
			     Turnovers_team),
	       Poss = possessions_basic(Two_Attempts + Three_Attempts,
					Freethrow_Attempts,
					Offensive_Rebounds,
					Turnovers),
	       GmSc = game_score(Points,
				 Twos + Threes,
				 Two_Attempts + Three_Attempts,
				 Freethrows,
				 Freethrow_Attempts,
				 Offensive_Rebounds,
				 Deffensive_Rebounds,
				 Assists,
				 Steals,
				 Blocks,
				 Turnovers,
				 Fouls),
	       pseudo_O_Rating = pseudo_offensive_rating(Twos,
							Threes,
							Freethrows,
							Assists,
							Offensive_Rebounds,
							Poss,
							Twos_team,
							Threes_team,
							Freethrows_team,
							Possessions_team),
	       D_Rating = defensive_rating(Seconds / 60,
					   Deffensive_Rebounds,
					   Steals,
					   Blocks,
					   Fouls,
					   MP_team,
					   Deffensive_Rebounds_team,
					   Steals_team,
					   Blocks_team,
					   Fouls_team,
					   Possessions_team,
					   Points_opponent,
					   MP_opponent,
					   Twos_opponent + Threes_opponent,
					   Two_Attempts_opponent + Three_Attempts_opponent,
					   Freethrows_opponent,
					   Freethrow_Attempts_opponent,
					   Offensive_Rebounds_opponent,
					   Deffensive_Rebounds_opponent,
					   Turnovers_opponent),
	       PProd = points_produced(Two_Attempts + Three_Attempts,
				       Freethrow_Attempts,
				       Turnovers,
				       pseudo_O_Rating),
	       PIE = player_impact_estimate(Points,
					    Twos + Threes,
					    Two_Attempts + Three_Attempts,
					    Freethrows,
					    Freethrow_Attempts,
					    Assists,
					    Steals,
					    Blocks,
					    Deffensive_Rebounds,
					    Offensive_Rebounds,
					    Turnovers,
					    Fouls,
					    Points_team + Points_opponent,
					    Twos_team + Threes_team +
					            Twos_opponent + Threes_opponent,
					    Two_Attempts_team + Three_Attempts_team +
					            Two_Attempts_opponent + Three_Attempts_opponent,
					    Freethrows_team + Freethrows_opponent,
					    Freethrow_Attempts_team + Freethrow_Attempts_opponent,
					    Assists_team + Assists_opponent,
					    Steals_team + Steals_opponent,
					    Blocks_team + Blocks_opponent,
					    Deffensive_Rebounds_team + Deffensive_Rebounds_opponent,
					    Offensive_Rebounds_team + Offensive_Rebounds_opponent,
					    Turnovers_team + Turnovers_opponent,
					    Fouls_team + Fouls_opponent)
	       ) %>%
	filter(Seconds > 0 | Injured == 1) %>%
	group_by(Player_ID, Season) %>%
	summarize(MP = sum(Seconds) / 60,
		  Avg_MP = mean(Seconds / 60),
		  Games = mean(Played),
		  Poss = sum(Poss),
		  FG = sum(Twos) + sum(Threes),
		  FGA = sum(Two_Attempts) + sum(Three_Attempts),
		  Twos = sum(Twos),
		  TwoPA = sum(Two_Attempts),
		  Threes = sum(Threes),
		  ThreePA = sum(Three_Attempts),
		  FT = sum(Freethrows),
		  FTA = sum(Freethrow_Attempts),
		  O_REB = sum(Offensive_Rebounds),
		  D_REB = sum(Deffensive_Rebounds),
		  AST = sum(Assists),
		  STL = sum(Steals),
		  BLK = sum(Blocks),
		  TOV = sum(Turnovers),
		  PF = sum(Fouls),
		  Pts = sum(Points),
		  Avg_Usage = mean(Usage),
		  Avg_GmSc = mean(GmSc),
		  Avg_O_Rating = mean(pseudo_O_Rating),
		  Avg_D_Rating = mean(D_Rating),
		  Inj_pct = mean(Injured),
		  PProd = sum(PProd),
		  PIE_sum = sum(PIE),
		  team_MP = sum(MP_team),
		  team_Poss = sum(Possessions_team),
		  team_Pace = mean(Pace_team),
		  team_FG = sum(Twos_team) + sum(Threes_team),
		  team_FGA = sum(Two_Attempts_team) + sum(Three_Attempts_team),
		  team_Twos = sum(Twos_team),
		  team_TwoPA = sum(Two_Attempts_team),
		  team_Threes = sum(Threes_team),
		  team_ThreePA = sum(Three_Attempts_team),
		  team_FT = sum(Freethrows_team),
		  team_FTA = sum(Freethrow_Attempts_team),
		  team_O_REB = sum(Offensive_Rebounds_team),
		  team_D_REB = sum(Deffensive_Rebounds_team),
		  team_AST = sum(Assists_team),
		  team_STL = sum(Steals_team),
		  team_BLK = sum(Blocks_team),
		  team_TO = sum(Turnovers_team),
		  team_PF = sum(Fouls_team),
		  team_Pts = sum(Points_team),
		  opp_MP = sum(MP_opponent),
		  opp_Poss = sum(Possessions_team),
		  opp_Pace = mean(Pace_opponent),
		  opp_FG = sum(Twos_opponent) + sum(Threes_opponent),
		  opp_FGA = sum(Two_Attempts_opponent) + sum(Three_Attempts_opponent),
		  opp_Twos = sum(Twos_opponent),
		  opp_TwoPA = sum(Two_Attempts_opponent),
		  opp_Threes = sum(Threes_opponent),
		  opp_ThreePA = sum(Three_Attempts_opponent),
		  opp_FT = sum(Freethrows_opponent),
		  opp_FTA = sum(Freethrow_Attempts_opponent),
		  opp_O_REB = sum(Offensive_Rebounds_opponent),
		  opp_D_REB = sum(Deffensive_Rebounds_opponent),
		  opp_AST = sum(Assists_opponent),
		  opp_STL = sum(Steals_opponent),
		  opp_BLK = sum(Blocks_opponent),
		  opp_TO = sum(Turnovers_opponent),
		  opp_PF = sum(Fouls_opponent),
		  opp_Pts = sum(Points_opponent),
		  .groups = 'keep'
	) %>%
	ungroup() %>%
	inner_join(league, by = c('Season')) %>%
	mutate(Touches = touches(FGA,
				 FTA,
				 AST,
				 TOV,
				 team_FTA,
				 opp_PF),
	       O_W_shares = offensive_win_shares(PProd,
						 Poss,
						 team_Pace,
						 league_Pace,
						 league_Pts_per_Poss,
						 league_Pts_per_game),
	       D_W_shares = defensive_win_shares(MP,
						 Avg_D_Rating,
						 team_MP,
						 team_Pace,
						 opp_Poss,
						 league_Pace,
						 league_Pts_per_Poss,
						 league_Pts_per_game),
	       win_shares = O_W_shares + D_W_shares
	) %>%
	inner_join(birthdays, by = c("Player_ID" = "Player_ID")) %>%
	inner_join(season_start, by = c("Season" = "Season")) %>%
	mutate(Age = floor(interval(Birthday, Start_Date) / years(1))) %>%
	filter(MP > 1000 | (Inj_pct > 0 & Avg_MP > 20)) %>%
	select(Player_ID,
	       Season,
	       MP,
	       Games,
	       Avg_MP,
	       Age,
	       Avg_Usage,
	       Touches,
	       team_Pace,
	       Inj_pct,
	       Avg_O_Rating,
	       Avg_D_Rating,
	       O_W_shares,
	       D_W_shares,
	       PIE_sum,
	       Avg_GmSc) %>%
	mutate(Two_back = Season - 1,
	       Three_back = Season - 2)

results <- player_season_totals %>%
	mutate(Next_Season_Games = Games) %>%
	select(Player_ID, Season, Next_Season_Games)

data <- player_season_totals %>%
	mutate(Season = Season + 1,
	       Age = Age + 1) %>%
	inner_join(player_season_totals %>%
		   	select(Player_ID,
			       Season,
			       Games,
			       MP,
			       Avg_MP,
			       Inj_pct,
			       PIE_sum,
			       Avg_GmSc),
		   by = c('Player_ID' = 'Player_ID',
			  'Two_back' = 'Season'),
		   suffix = c('', '_two_back')) %>%
	inner_join(player_season_totals %>%
		   	select(Player_ID,
			       Season,
			       Games,
			       MP,
			       Avg_MP,
			       Inj_pct,
			       PIE_sum,
			       Avg_GmSc),
		   by = c('Player_ID' = 'Player_ID',
			  'Three_back' = 'Season'),
		   suffix = c('', '_three_back')) %>%
	inner_join(player_season_totals %>%
		mutate(Next_Season_Games = Games) %>%
		select(Player_ID,
		       Season,
		       Next_Season_Games),
		       by = c('Player_ID' = 'Player_ID',
			      'Season' = 'Season')) %>%
	select(!c('Season',
		  'Two_back',
		  'Three_back')) %>%
	group_by(Player_ID)

data$Player_ID <- as_factor(data$Player_ID)
```

As it stands, this only leaves me with `r dim(data)[0]` data points.
The aggregation and joining of data spanning multiple years tremendously reduces the data available,
but it is still a very large number given the subjects of interest in a given year might only add up to `r 12*13`.

## Data explanation

I will assume a basic understanding of advanced basketball statistics,
but for the most part, even that's not needed for most of the data that will likely be useful.
Keep in mind too that if not otherwise stated,
the explanation is referring to the data a season prior to the season of interest.
Here is a simple explanation of each piece of data I have available to me for this analysis:

| Variable            | Explanation                                                                     |
|---------------------|---------------------------------------------------------------------------------|
| Player_ID           | Id number given to each player                                                  |
| MP                  | Total minutes played                                                            |
| Games               | Proportion of games played                                                      |
| Avg_MP              | Average number of minutes played in each game that was played                   |
| Age                 | Age at the start of the season                                                  |
| Avg_Usage           | Average usage rate                                                              |
| Touches             | Number of touches a player got in a season                                      |
| team_Pace           | The average pace for the player's team (following the player if traded)         |
| Inj_pct             | Proportion of games where the player was injured                                |
| Avg_O_Rating        | Average offensive rating, using my basic approximation                          |
| Avg_D_Rating        | Average defensive rating                                                        |
| O_W_shares          | Total offensive win shares, based off my approximation of offensive rating      |
| D_W_shares          | Total defensive win shares                                                      |
| PIE_sum             | Sum of Player Impact Estimate for each game                                     |
| Avg_GmSc            | Average player game score                                                       |
| Games_two_back      | Proportion of games played two seasons ago                                      |
| MP_two_back         | Total minutes played two seasons ago                                            |
| Avg_MP_two_back     | Average number of minutes played in each game that was played two seasons ago   |
| Inj_pct_two_back    | Proportion of games where the player was injured two seasons ago                |
| PIE_sum_two_back    | Sum of Player Impact Estimate for each game two seasons ago                     |
| Avg_GmSc_two_back   | Average player game score two seasons ago                                       |
| Games_three_back    | Proportion of games played three seasons ago                                    |
| MP_three_back       | Total minutes played three seasons ago                                          |
| Avg_MP_three_back   | Average number of minutes played in each game that was played three seasons ago |
| Inj_pct_three_back  | Proportion of games where the player was injured three seasons ago              |
| PIE_sum_three_back  | Sum of Player Impact Estimate for each game three seasons ago                   |
| Avg_GmSc_three_back | Average player game score three seasons ago                                     |
| Next_Season_Games   | Proportion of games played in the season of interest                            |

# Assumptions

```{r linearity}

```

```{r residuals}

```

```{r multicollinearity}

```
