---
title: "Modeling NBA Number of Games Played by Players"
author: "Carson Buttars"
date: ""
output: pdf_document
---

```{r setup, include=FALSE}
library(DBI)
library(tidyverse)
library(lubridate)
library(devtools)
load_all('lib/advanced.stats/')
conn = dbConnect(RSQLite::SQLite(), "../data/bball_db")

birthdays <- dbReadTable(conn, "player_info") %>%
	as_tibble() %>%
	mutate(Birthday = mdy(Birthday)) %>%
	select(Player_ID, Birthday)

season_start <- dbReadTable(conn, "game_info") %>%
	as_tibble() %>%
	mutate(Date = mdy(Date)) %>%
	group_by(Season) %>%
	summarize(Start_Date = first(Date)) %>%
	select(Season, Start_Date)

team_games <- dbReadTable(conn, "team_games") %>%
	rename(MP = Total_minutes) %>%
	as_tibble()

team_and_opponents <- team_games %>%
	inner_join(team_games, by = c("Game_ID" = "Game_ID",
				      "Opponent_ID" = "Team_ID",
				      "Team_ID" = "Opponent_ID",
				      "Season" = "Season"),
		   suffix = c("_team", "_opponent")) %>%
	mutate(Pace_team = pace(MP_team, Possessions_team, Possessions_opponent),
	       Pace_opponent = pace(MP_opponent, Possessions_opponent, Possessions_team))

league <- team_and_opponents %>%
	group_by(Season) %>%
	summarize(league_Games = n(),
		  league_MP = sum(MP_team),
		  league_Possessions_team = sum(Possessions_team),
		  league_Threes_team = sum(Threes_team),
		  league_Three_Attempts_team = sum(Three_Attempts_team),
		  league_Twos_team = sum(Twos_team),
		  league_Two_Attempts_team = sum(Two_Attempts_team),
		  league_Freethrows_team = sum(Freethrows_team),
		  league_Freethrow_Attempts_team = sum(Freethrow_Attempts_team),
		  league_Offensive_Rebounds_team = sum(Offensive_Rebounds_team),
		  league_Deffensive_Rebounds_team = sum(Deffensive_Rebounds_team),
		  league_Assists_team = sum(Assists_team),
		  league_Steals_team = sum(Steals_team),
		  league_Blocks_team = sum(Blocks_team),
		  league_Turnovers_team = sum(Turnovers_team),
		  league_Fouls_team = sum(Fouls_team),
		  league_Points_team = sum(Points_team),
		  league_Win_team = mean(Win_team),
		  league_Home_team = mean(Home_team),
		  league_MP = sum(MP_opponent),
		  league_Possessions_opponent = sum(Possessions_opponent),
		  league_Threes_opponent = sum(Threes_opponent),
		  league_Three_Attempts_opponent = sum(Three_Attempts_opponent),
		  league_Twos_opponent = sum(Twos_opponent),
		  league_Two_Attempts_opponent = sum(Two_Attempts_opponent),
		  league_Freethrows_opponent = sum(Freethrows_opponent),
		  league_Freethrow_Attempts_opponent = sum(Freethrow_Attempts_opponent),
		  league_Offensive_Rebounds_opponent = sum(Offensive_Rebounds_opponent),
		  league_Deffensive_Rebounds_opponent = sum(Deffensive_Rebounds_opponent),
		  league_Assists_opponent = sum(Assists_opponent),
		  league_Steals_opponent = sum(Steals_opponent),
		  league_Blocks_opponent = sum(Blocks_opponent),
		  league_Turnovers_opponent = sum(Turnovers_opponent),
		  league_Fouls_opponent = sum(Fouls_opponent),
		  league_Points_opponent = sum(Points_opponent),
		  league_Win_opponent = mean(Win_opponent),
		  league_Home_opponent = mean(Home_opponent),
		  league_Pts_per_game = mean(Points_team)
	) %>%
	mutate(league_Pace = pace(league_MP, league_Possessions_team, league_Possessions_opponent),
	       league_Pts_per_Poss = league_Points_team / league_Possessions_team)

player_season_totals <- dbReadTable(conn, "player_games") %>%
	as_tibble() %>%
	inner_join(team_and_opponents, by = c("Game_ID" = "Game_ID",
					      "Team_ID" = "Team_ID",
					      "Opponent_ID" = "Opponent_ID",
					      "Season" = "Season")) %>%
	mutate(Poss = possessions_basic(Two_Attempts + Three_Attempts,
					Freethrow_Attempts,
					Offensive_Rebounds,
					Turnovers),
	       GmSc = game_score(Points,
				 Twos + Threes,
				 Two_Attempts + Three_Attempts,
				 Freethrows,
				 Freethrow_Attempts,
				 Offensive_Rebounds,
				 Deffensive_Rebounds,
				 Assists,
				 Steals,
				 Blocks,
				 Turnovers,
				 Fouls),
	       O_Rating = offensive_rating(Points,
					   Seconds / 60,
					   Two_Attempts + Three_Attempts,
					   Twos + Threes,
					   Threes,
					   ThreePA,
					   Freethrow_Attempts,
					   Freethrows,
					   Assists,
					   Offensive_Rebounds,
					   Turnovers,
					   Points_team,
					   MP_team,
					   Twos_team + Threes_team,
					   Two_Attempts_team + Three_Attempts_team,
					   Threes_team,
					   ThreePA_team,
					   Freethrows_team,
					   Freethrow_Attempts_team,
					   Assists_team,
					   Offensive_Rebounds_team,
					   Turnovers_team,
					   Offensive_Rebounds_opponent + Deffensive_Rebounds_opponent,
					   Offensive_Rebounds_opponent),
	       D_Rating = defensive_rating(Seconds / 60,
					   Deffensive_Rebounds,
					   Steals,
					   Blocks,
					   Fouls,
					   MP_team,
					   Deffensive_Rebounds_team,
					   Steals_team,
					   Blocks_team,
					   Fouls_team,
					   Possessions_team,
					   Points_opponent,
					   MP_opponent,
					   Twos_opponent + Threes_opponent,
					   Two_Attempts_opponent + Three_Attempts_opponent,
					   Freethrows_opponent,
					   Freethrow_Attempts_opponent,
					   Offensive_Rebounds_opponent,
					   Deffensive_Rebounds_opponent,
					   Turnovers_opponent),
	       PProd = points_produced(Two_Attempts + Three_Attempts,
				       Freethrow_Attempts,
				       Turnovers,
				       O_Rating),
	       PIE = player_impact_estimate(Points,
					    Twos + Threes,
					    Two_Attempts + Three_Attempts,
					    Freethrows,
					    Freethrow_Attempts,
					    Assists,
					    Steals,
					    Blocks,
					    Deffensive_Rebounds,
					    Offensive_Rebounds,
					    Turnovers,
					    Fouls,
					    Points_team + Points_opponent,
					    Twos_team + Threes_team +
					            Twos_opponent + Threes_opponent,
					    Two_Attempts_team + Three_Attempts_team +
					            Two_Attempts_opponent + Three_Attempts_opponent,
					    Freethrows_team + Freethrows_opponent,
					    Freethrow_Attempts_team + Freethrow_Attempts_opponent,
					    Assists_team + Assists_opponent,
					    Steals_team + Steals_opponent,
					    Blocks_team + Blocks_opponent,
					    Deffensive_Rebounds_team + Deffensive_Rebounds_opponent,
					    Offensive_Rebounds_team + Offensive_Rebounds_opponent,
					    Turnovers_team + Turnovers_opponent,
					    Fouls_team + Fouls_opponent)
	       ) %>%
	group_by(Player_ID, Season) %>%
	summarize(MP = sum(Seconds) / 60,
		  Games = n(),
		  Poss = sum(Poss),
		  FG = sum(Twos) + sum(Threes),
		  FGA = sum(Two_Attempts) + sum(Three_Attempts),
		  Twos = sum(Twos),
		  TwoPA = sum(Two_Attempts),
		  Threes = sum(Threes),
		  ThreePA = sum(Three_Attempts),
		  FT = sum(Freethrows),
		  FTA = sum(Freethrow_Attempts),
		  O_REB = sum(Offensive_Rebounds),
		  D_REB = sum(Deffensive_Rebounds),
		  AST = sum(Assists),
		  STL = sum(Steals),
		  BLK = sum(Blocks),
		  TO = sum(Turnovers),
		  PF = sum(Fouls),
		  Pts = sum(Points),
		  Avg_GmSc = mean(GmSc),
		  Avg_O_Rating = mean(O_Rating),
		  Avg_D_Rating = mean(D_Rating),
		  Inj_pct = mean(Injured),
		  PProd = sum(PProd),
		  PIE_sum = sum(PIE),
		  team_MP = sum(MP_team),
		  team_Poss = sum(Possessions_team),
		  team_Pace = mean(Pace_team),
		  team_FG = sum(Twos_team) + sum(Threes_team),
		  team_FGA = sum(Two_Attempts_team) + sum(Three_Attempts_team),
		  team_Twos = sum(Twos_team),
		  team_TwoPA = sum(Two_Attempts_team),
		  team_Threes = sum(Threes_team),
		  team_ThreePA = sum(Three_Attempts_team),
		  team_FT = sum(Freethrows_team),
		  team_FTA = sum(Freethrow_Attempts_team),
		  team_O_REB = sum(Offensive_Rebounds_team),
		  team_D_REB = sum(Deffensive_Rebounds_team),
		  team_AST = sum(Assists_team),
		  team_STL = sum(Steals_team),
		  team_BLK = sum(Blocks_team),
		  team_TO = sum(Turnovers_team),
		  team_PF = sum(Fouls_team),
		  team_Pts = sum(Points_team),
		  opp_MP = sum(MP_opponent),
		  opp_Poss = sum(Possessions_team),
		  opp_Pace = mean(Pace_opponent),
		  opp_FG = sum(Twos_opponent) + sum(Threes_opponent),
		  opp_FGA = sum(Two_Attempts_opponent) + sum(Three_Attempts_opponent),
		  opp_Twos = sum(Twos_opponent),
		  opp_TwoPA = sum(Two_Attempts_opponent),
		  opp_Threes = sum(Threes_opponent),
		  opp_ThreePA = sum(Three_Attempts_opponent),
		  opp_FT = sum(Freethrows_opponent),
		  opp_FTA = sum(Freethrow_Attempts_opponent),
		  opp_O_REB = sum(Offensive_Rebounds_opponent),
		  opp_D_REB = sum(Deffensive_Rebounds_opponent),
		  opp_AST = sum(Assists_opponent),
		  opp_STL = sum(Steals_opponent),
		  opp_BLK = sum(Blocks_opponent),
		  opp_TO = sum(Turnovers_opponent),
		  opp_PF = sum(Fouls_opponent),
		  opp_Pts = sum(Points_opponent),
		  .groups = 'keep'
	) %>%
	inner_join(league, by = c('Season')) %>%
	mutate(Usage = usage(MP, FGA, FTA, TO, team_MP, team_FGA, team_FTA, team_TO),
	       O_W_shares = offensive_win_shares(PProd,
						 Poss,
						 team_Pace,
						 league_Pace,
						 league_Pts_per_Poss,
						 league_Pts_per_game),
	       D_W_shares = defensive_win_shares(MP,
						 Avg_D_Rating,
						 team_MP,
						 team_Pace,
						 opp_Poss,
						 league_Pace,
						 league_Pts_per_Poss,
						 league_Pts_per_game),
	       win_shares = O_W_shares + D_W_shares
	) %>%
	inner_join(birthdays, by = c("Player_ID" = "Player_ID")) %>%
	inner_join(season_start, by = c("Season" = "Season")) %>%
	mutate(Age = floor(interval(Birthday, Start_Date) / years(1)))

```

---
# The response variable: games per season
# Explanatory variables I want to grab:
# - Age (at beginning of season)
# - usage
# - Last year's touches
# - Last year's team pace
# - Last year's minutes per game
# - Last 2 years' minutes per game
# - Last 3 years' games played %
# - Last 5 years' games played %
# - Career proportion of games injured
# - Offensive Rating
# - Defensive Rating
# - Offensive Win Shares
# - Defensive Win Shares
# - Player Impact Estimate
# - Average Game Score
---

```{r data_cleanup, include=FALSE}

```
