---
title: "Modeling NBA Number of Games Played by Players"
author: "Carson Buttars"
date: ""
output: pdf_document
---

```{r setup, include=FALSE}
library(DBI)
library(tidyverse)
conn = dbConnect(RSQLite::SQLite(), '../data/bball_db')

player_info <- dbReadTable(conn, 'player_info') %>%
	as_tibble() %>%
	mutate(Birthday = mdy(Birthday),
		Debut_Date = mdy(Debut_Date)
	)
player_season_totals <- dbReadTable(conn, 'player_games') %>%
	as_tibble() %>%
	group_by(Player_ID, Team_ID) %>%
	summarize(MP = sum(Seconds) / 60,
		Games = n(),
		FG = sum(Twos) + sum(Threes),
		FGA = sum(Two_Attempts) + sum(Three_Attempts),
		Twos = sum(Twos),
		TwoPA = sum(Two_Attempts),
		Threes = sum(Threes),
		ThreePA = sum(Three_Attempts),
		FT = sum(Freethrows),
		FTA = sum(Freethrow_Attempts),
	)
```

---
# The response variable: games per season
# Explanatory variables I want to grab:
# - Age (at beginning of season)
# - usage
# - Last year's touches
# - Last year's team pace
# - Last year's minutes per game
# - Last 2 years' minutes per game
# - Last 3 years' games played %
# - Last 5 years' games played %
# - Career proportion of games injured
# - Offensive Rating
# - Defensive Rating
# - Offensive Win Shares
# - Defensive Win Shares
# - Player Impact Estimate
# - Average Game Score
---

```{r data_cleanup, include=FALSE}

```
